---
title: "Exploring response and covariates data for seafood production model"
author: "Marina Espinasse"
date: "`r Sys.Date()`"
output:
  html_document:
    css: '~/github/spfood/src/style.css'
    toc: true
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/spfood/src/spfood_banner.html'
  pdf_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
source('~/github/spfood/src/common.R')
source('~/github/spfood/src/HighstatLibV13.R')
```

Here I will run the exploration of the response variables (aquaculture production and fisheries production) and of the covariates

```{r}
library(vroom)
library(here)
library(lattice)
library(cowplot)
```
# Load the data
```{r}
seafood <-read.csv(here("prep", "output_data", "spfood_allvars.csv"))
```

Calculate total sea food production here as well. I also create a log-transformed fisheries catch, because it is highly skewed to the left (see histograms below)

```{r}
seafood_prep <- seafood |> 
  rowwise() |> 
  mutate(totprod = sum(aqua_prod_ton, total_catch_ton, na.rm=T)) |> 
  mutate(aqua_prop = aqua_prod_ton/totprod) |> 
  mutate(fish_catch_log = log(total_catch_ton + 1)) |> 
  mutate(aqua_prod_log = log(aqua_prod_ton + 1))
```



# Data exploration
I will use extensively the approaches and functions from the two Highland statistics ltd books:
*  Mixed effects models and extensions in ecology with R. (2009).
Zuur, AF, Ieno, EN, Walker, N, Saveliev, AA, and Smith, GM. Springer
 
and 

*  #Library files for courses provided by: Highland Statistics Ltd.
Zuur, Alain F., and Elena N. Ieno. "Spatial, temporal and spatial-temporal ecological data analysis with r-inla. Volume II" Highland Statistics Ltd 1 (2019).

Cleveland dotplots of the variables
```{r}
Mydotplot(seafood_prep[,c(7,8,9,10,11,12,13,14,15)])
```


We observe an excess of zeros or close-to-zero values in the population, population growth, unemployed, and total catch variables.

Check the % of zeroes in the variables:

```{r}
zeros_prop <- function(x){ #x is a numerical vector
  rows = sum(!is.na(x)) 
  zeros = sum(x == 0, na.rm = T)
  prop = zeros/rows
  prop
}
```


```{r}
myvars <- seafood_prep[,c("aqua_prod_ton",
                          "total_catch_ton",
                          "population",
                          "popgrowth",
                          "sea_area",
                          "unemployed",
                          "wageco")]
map(myvars, ~zeros_prop(.x))
```

Only fisheries has 15% of zeros, other variables do not suffer from zero inflation

Check out the distribution of variables:
```{r, results='hide',fig.keep='all'}
varnames <- list("aquaculture", "fisheries", "population", "population growth",
                 "sea area",
                 "unemployed",
                 "p90/p10")
par(mfrow = c(2,4))
hist <- map2(myvars,varnames, ~ hist(.x, main = .y, col = "seagreen", xlab = " ", ylab = ""))
```

## Historgrams of the 59 selected municipalities
For the mixed models, I will use only 59 municipalities, that is, 817 out of 862 observations (that had at least 10 years of data). Let's see if historgamrs are differenct for these observations.

```{r}
mcp59 <- read_csv(here::here("prep", "output_data", "mcp_59selected.csv"))

myvars59 <- mcp59[,c("aqua_prod_ton",
                          "total_catch_ton",
                          "population",
                          "popgrowth",
                          "sea_area",
                          "unemployed",
                          "wageco")]
par(mfrow = c(2,4))
hist <- map2(myvars59,varnames, ~ hist(.x, main = .y, col = "seagreen", xlab = " ", ylab = ""))
```


Most of the variables are left-skewed, but fisheries may need to be transformed, or later, the response variable aquaculture/(aquaculture + fisheries) may need to be transformed.

The boxplots of the variables, to check if there are outliers (formally): yes, since the variables are all left-skewed
```{r, results='hide',fig.keep='all'}
par(mfrow = c(2,4))
map2(myvars,varnames, ~ boxplot(.x, main = .y, col = "seagreen", xlab = " "))
```


Let's take a look at the variable aqua_prop which is a proportion of aquaculture in the total seafood production and at the log of the fisheries catches
```{r}
par(mfrow=c(1,3))
hist(seafood_prep$aqua_prop, main = "Aquaculture proportion", xlab = "", col = "orchid")
hist(seafood_prep$fish_catch_log, main = "Log of fisheries catches", xlab = "", col = "orchid")
hist(seafood_prep$aqua_prod_log, main = "Log of aquaculture production", xlab = "", col = "orchid")
```



# Collinearity
```{r}
Mypairs(myvars)
```


Clear association is only seen for the population and growth, and between the number of unemployed and population growth (positive relationship in both cases).

# Responses versus covariates

##  Aquaculture proportion versus covariates
```{r}
MyX <- c("year", "sea_area", "wageco", "population", "popgrowth" , "unemployed")
MyMultipanel.ggp2(Z = seafood_prep,
                  varx = MyX,
                  vary = "aqua_prop",
                  ylab = "aquaculture % in seafood production",
                  addSmoother = TRUE,
                  addRegressionLine = FALSE,
                  addHorizontalLine = FALSE) 

```

No clear associations or patterns, but i need to consider transformation of a response variable.

## Aquaculture production versus covariates
What if aquaculture production alone would be used as a response variable? Let's see also if fisheries catch has anything to do with aquaculture produciton


```{r}

MyMultipanel.ggp2(Z = seafood_prep,
                  varx = MyX,
                  vary = "aqua_prod_ton",
                  ylab = "aquaculture production",
                  addSmoother = TRUE,
                  addRegressionLine = FALSE,
                  addHorizontalLine = FALSE) 
```


This looks more interesting! Relationships are close to linear, but for population and wage coefficient maybe not.

## Fisheries catch versus aquaculture

Might be interesting to examine their associations statistically
```{r}
fish_aqua <-ggplot(
  data = seafood_prep,
  mapping = aes(x = total_catch_ton, y = aqua_prod_ton)) +
  geom_point() +
  geom_smooth()

fish_aqua
```

## Fisheries production versus covariates
```{r}

MyMultipanel.ggp2(Z = seafood_prep,
                  varx = MyX,
                  vary = "total_catch_ton",
                  ylab = "fisheries catch",
                  addSmoother = TRUE,
                  addRegressionLine = FALSE,
                  addHorizontalLine = FALSE) 
```

I maybe will need to transform fisheries catch to log(catch) for the analysis


# Checking the distribution of responce variables per municipality

Just to see if the distribution of fisheries and aquaculture production over years look very different when plotted per municipality.
There is a clear increase in the production over years!
```{r, fig.width=10, fig.height=7, warning = F}
ggplot(data = seafood_prep) +
  geom_point(mapping = aes(x = year, y =  aqua_prod_ton), size = 0.5) +
  geom_smooth(mapping = aes(x = year, y =  aqua_prod_ton), size = 0.4) +
  facet_wrap( municip ~ .) +
  theme(
    axis.text.x = element_text(size = 10, angle = 90),
    legend.position = "none",
    panel.background = element_rect(fill = "white")) 
```


Same but for fisheries
```{r, fig.width=10, fig.height=7, warning = F}
ggplot(data = seafood_prep) +
  geom_point(mapping = aes(x = year, y =  total_catch_ton), size = 0.5) +
  geom_smooth(mapping = aes(x = year, y =  total_catch_ton), size = 0.4) +
  facet_wrap( municip ~ .) +
  theme(
    axis.text.x = element_text(size = 10, angle = 90),
    legend.position = "none",
    panel.background = element_rect(fill = "white"))
  
```


# Verifying dependence in the response variables
I am already quite sure that both responses will be temporally and spatially dependent. But to test for that formally, we can use an autocorrelation test:

```{r}
seafood_temp <- arrange(seafood_prep, year) 
  acf(seafood_temp$aqua_prod_ton, na.action = na.pass)

```

```{r}
acf(seafood_temp$total_catch_ton)
```


