---
title: "Mapping fisheries  production"
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    css: '~/github/nor-prep/prep/templates/style.css'
    toc: true
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/nor-prep/prep/templates/norway_banner.html'
  pdf_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
    echo = TRUE, 
    out.width = "100%",
    dpi = 300,
    fig.width = 8,
    fig.height = 6,
    strip.white = T)
source('~/github/spfood/src/common.r')
```

```{r, message=F}
library(mapdata)
library(maptools)
library(rgdal)
library(rgeos)
library(maps)
library(rnaturalearth)
library(rnaturalearthdata)
library(sf)
library(here)
```


# Loading data
## Municipalities shape files
```{r}
allkommuner <-readOGR(dsn = "/Users/marinaespinasse/github/nor-prep/prep/administrative/raw/shapefiles/NO_AdminOmrader")
```

# Fisheries annual production data

Loading production data  and also Norwegian names of municipalities (with special Norwegian characters).
For the spfood paper I need only the municipalities that both had fisheries and aquaculture and that have at least 10 years of observations. I load the csv table with the 59 municipalities, 

```{r}
komnames_norw <- read_excel(file.path("./raw_data/komlist_norw.xlsx"))
seafood <- read_csv(here("prep", "output_data", "mcp_59selected.csv"))
```

# Preparing mapping layers 

Change the municipality number of Harstad to 1901, as it was before

```{r}
seafood_prep <- seafood %>% 
  select(-totobs) |> 
  mutate(Komnum = replace(municip_number,
                          municip_number == 1903,
                          1901))

seafood_shapes <-allkommuner[allkommuner$KOMM %in% seafood_prep$Komnum,] 
```


```{r }
world <- ne_countries(scale = "medium", returnclass = "sf")

kommuner.pol <- spTransform(seafood_shapes, CRS("+proj=longlat +ellps=GRS80"))
kommunersf <- sf::st_as_sf(seafood_shapes, coords= c("long", "lat"),crs = 4326, agr="identity")

kommunersf_prep <- kommunersf %>% 
  mutate(KOMM = replace(KOMM,
                        KOMM == 1901,
                        1903))
```
  

## Preparing the data for year 2005

```{r }
fishcatch05 <-filter(seafood, year == "2005")

plot_fishcatch05_prep <- fishcatch05 %>% 
  left_join(
  kommunersf_prep, 
  by = c("municip_number"="KOMM")) %>% 
  st_as_sf(.)
```



## Preparing the data for year 2018
I do all the same manipulations for catch in year 2018 as for year 2005. I calculate total catch of all species,
and then join the catches table with a geometry of municipalities.
```{r }
fishcatch18 <-filter(seafood, year == "2018")

plot_fishcatch18_prep <-  fishcatch18 %>% 
  left_join(
  kommunersf_prep, 
  by = c("municip_number"="KOMM")) %>% 
  st_as_sf(.)
```


# Mapping fisheries production
## 2005 maps

It is better that both maps has similar number of intervals for legend, like 6.
I also divide the catches by 1000 to have tons on the map.

```{r}

intervals05 <-c(0, 8000, 16000, 24000, 32000,40000,46000)#intervals for year 2005 only

labels05 <- imap_chr(intervals05, function(., id){
  return(paste0(formatC(intervals05[id], format="f", big.mark="," , digits = 0),
                             "",
                             " – ",
                             formatC(intervals05[id + 1], format="f", big.mark="," , digits = 0),
                             ""))
})

labels05 <- labels05[1:length(labels05) - 1]


plot_fishcatch05_prep2  <- plot_fishcatch05_prep %>% 
  filter(!is.na(total_catch_ton)) %>% 
  mutate(catch_quantiles = cut(total_catch_ton,
                               breaks = intervals05,
                               labels = labels05,
                               include.lowest = T)) 


```

For adding municipalities names, we need centroids of each municipality's polygon.
I will also remove the municipalities 1920 and 1919 because they are too small, and numbers don't fit, i'll add number manually afterwards.
```{r}
komnames05 <- cbind(plot_fishcatch05_prep2,st_coordinates(st_centroid(plot_fishcatch05_prep2)))
komnames_prep05 <- filter(komnames05, !municip_number %in% c(1919, 1920))
```

Add transparency to map colors, as suggested here: https://gotellilab.github.io/GotelliLabMeetingHacks/NickGotelli/ColorPalettes.html

```{r}

addalpha <- function(colors, alpha=0.7) {
  r <- col2rgb(colors, alpha=T)
  # Apply alpha
  r[4,] <- alpha*255
  r <- r/255.0
  return(rgb(r[1,], r[2,], r[3,], r[4,]))
}

```

```{r}
cols <-addalpha(brewer.pal(6,"Greens"))
```


```{r}
base_map05 <- ggplot() +
  geom_sf(
   data = world, 
   fill = "antiquewhite1", 
   color = "dimgray", 
   size = 0.3) +
   #coord_sf(xlim = c(10, 32), ylim = c(65, 72)
  theme(axis.text = element_text(size = 17),
        legend.text = element_text(size = 15),
        legend.title = element_text(size = 15,  face = "bold"),
        panel.background = element_rect(fill = "aliceblue") ,
        panel.grid.major = element_line(color = "#dbdbd9", linetype = "dashed"),
        plot.title = element_text(hjust = 0.5, vjust = 0.5, size = 17, face = "bold")
  ) +
  geom_sf(
    data = plot_fishcatch05_prep2,
    mapping = aes(fill = catch_quantiles),
    color = "white",
    size = 0.2
  ) +
  
  coord_sf(xlim = c(8, 32), ylim = c(64, 72), expand = FALSE) +
  labs(x = "Longitude", y = "Latitude") +
  scale_fill_manual(
    values = cols,
    name = "Fisheries catch ranges (ton)",
    guide = guide_legend(
      keyheight = unit(8, units = "mm"),
      title.position = "top"
    )
  ) +
  labs(
    x = NULL,
    y = NULL,
    title = "2005"
  ) 

```

```{r, dpi = 250, fig.height=8, fig.width=8}
main_map05 <- base_map05 +
  geom_path(kommuner.pol, mapping = aes(x = long, y = lat, group = id), color = 'dimgray',  size = .4) +    geom_sf_text(data = komnames_prep05, aes(X,Y, label= municip_number),fontface = "bold", size = 2) 
main_map05
```



## 2018 map 
Similarly, prepare municipalities numbers - only for those, that had registered catches in 2018.
```{r}
komnames18 <- cbind(plot_fishcatch18_prep,st_coordinates(st_centroid(plot_fishcatch18_prep)))
komnames_prep18 <- filter(komnames18, !municip_number %in% c(1919, 1920))
```

```{r}

intervals18 <-c(0, 8000, 16000, 24000, 32000)#intervals for year 2018 only

labels18 <- imap_chr(intervals18, function(., id){
  return(paste0(formatC(intervals18[id], format="f", big.mark="," , digits = 0),
                             "",
                             " – ",
                             formatC(intervals18[id + 1], format="f", big.mark="," , digits = 0),
                             ""))
})

labels18 <- labels18[1:length(labels18) - 1]

plot_fishcatch18_prep2  <- plot_fishcatch18_prep %>% 
  filter(!is.na(total_catch_ton)) %>% 
  mutate(catch_quantiles = cut(total_catch_ton,
                               breaks = intervals18,
                               labels = labels18,
                               include.lowest = T))
```

```{r}
base_map18 <- ggplot() +
  geom_sf(
   data = world, 
   fill = "antiquewhite1", 
   color = "dimgray", 
   size = 0.3) +
   #coord_sf(xlim = c(10, 32), ylim = c(65, 72)
  theme(axis.text = element_text(size = 17),
        legend.text = element_text(size = 15),
        legend.title = element_text(size = 15,  face = "bold"),
        panel.background = element_rect(fill = "aliceblue") ,
        panel.grid.major = element_line(color = "#dbdbd9", linetype = "dashed"),
        plot.title = element_text(hjust = 0.5, vjust = 0.5, size = 17, face = "bold")
  ) +
  geom_sf(
    data = plot_fishcatch18_prep2,
    mapping = aes(fill = catch_quantiles),
    color = "white",
    size = 0.2,
    show.legend = FALSE
  ) +
  
  coord_sf(xlim = c(8, 32), ylim = c(64, 72), expand = FALSE) +
  labs(x = "Longitude", y = "Latitude") +
  scale_fill_manual(
    values = cols
    # name = "Fisheries catch ranges, tons",
    # guide = guide_legend(
    #   keyheight = unit(8, units = "mm"),
    #   title.position = "none"
    ) +
  labs(
    x = NULL,
    y = NULL,
    title = "2018"
  ) 

```

```{r, dpi = 250, fig.height=8, fig.width=8}
main_map18 <- base_map18 +geom_path(kommuner.pol, mapping = aes(x = long, y = lat, group = id), color = 'dimgray',  size = .4) +geom_sf_text(data = komnames_prep18, aes(X,Y, label= municip_number),fontface = "bold", size = 2) 
main_map18
```

# Aligning both plots together
```{r}
library(patchwork)
```

```{r fig.height=5, fig.width=4, dpi = 400}
grDevices::cairo_pdf("./figs/fisheries_catch_2005_2018.pdf", width = 10, height = 12, fallback_resolution = 300)
main_map05/main_map18
```

# Plot all years together

I will include this figure in the supplementary material

Make seafood dataframe an sf object
```{r}
plot_fishcatch_all <-  seafood %>% 
  # mutate(municip_number = as.double(municip_number)) %>%
  # filter(municip_number %in% spfoodkom_prep$municip_number) |> 
  left_join(
  kommunersf_prep, 
  by = c("municip_number"="KOMM")) %>% 
  st_as_sf(.)
```

Create intervals of total catch for all years, in the same way as for 2005 and 2018.
```{r}
intervals <-c(0, 8000, 16000, 24000, 32000,40000, 48000, 56000, 61000) 
# 
labels_all <- imap_chr(intervals, function(., id){
  return(paste0(formatC(intervals[id], format="f", big.mark="," , digits = 0),
                             "",
                             " – ",
                             formatC(intervals[id + 1], format="f", big.mark="," ,                                digits = 0),
                             ""))
})

labels_all <- labels_all[1:length(labels_all) - 1]

plot_fishcatch_all2  <- plot_fishcatch_all %>% 
  filter(!is.na(total_catch_ton)) %>%
  mutate(catch_quantiles = cut(total_catch_ton,
                               breaks = intervals,
                               labels = labels_all,
                               include.lowest = T))
```

```{r}
cols2 <-brewer.pal(8,"Greens")
```



```{r}
base_map_all <- ggplot() +
  geom_sf(
   data = world, 
   fill = "antiquewhite1", 
   color = "dimgray", 
   size = 0.3) +
  theme(axis.text = element_text(size = 15),
        strip.text = element_text(size = 15, face = "bold"),
        legend.text = element_text(size = 15),
        legend.title = element_text(size = 15,  face = "bold"),
        panel.background = element_rect(fill = "aliceblue") ,
        strip.background = element_rect(fill ="gray96"),
        panel.grid.major = element_line(color = "#dbdbd9", linetype = "dashed"),
        plot.title = element_text(hjust = 0.5, vjust = 0.5, size = 17, face = "bold")
  ) +
  geom_sf(
    data = plot_fishcatch_all2,
    mapping = aes(fill = catch_quantiles),
    color = "white",
    size = 0.2
  ) +
  coord_sf(xlim = c(8, 32), ylim = c(64, 72), expand = FALSE) +
  labs(x = "Longitude", y = "Latitude") +
  scale_fill_manual(
    values = cols2,
    name = "Fisheries catch ranges (ton)",
    guide = guide_legend(
      keyheight = unit(8, units = "mm"),
      title.position = "top")) +
  labs(
    x = NULL,
    y = NULL,
    title = ""
  ) +
  facet_wrap(year ~.)
```

```{r}
main_map_all <- base_map_all +
  geom_path(kommuner.pol, mapping = aes(x = long, y = lat, group = id), color = 'dimgray',  size = .2) 
```


```{r}
main_map_all
```

```{r, fig.width=10, fig.height=7}
 

ggsave("./figs/fish_catch_all_years.tiff", main_map_all, width = 18, height = 16, dpi = 300)

```

