---
title: "Creating maps of municipalities and counties of northern Norway"
author: "Marina Espinasse"
date: "`r Sys.Date()`"
output:
  html_document:
    css: '~/github/spfood/src/style.css'
    code_folding: show
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: true
    theme: lumen
    highlight: espresso
  pdf_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set
source('~/github/spfood/src/common.r')
```

Loading mapping libraries
```{r include = FALSE}
library(mapdata)
library(maptools)
library(rgdal)
library(maps)
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)
```

# Loading shapefiles of municipalities, counties, and the list of municipalities of the project

Many of the files are taken from another open repository - nor-prep
```{r}
fylker <-readOGR(dsn = "/Users/marinaespinasse/github/nor-prep/prep/administrative/raw/shapefiles/NO_Fylker")
allkommuner <-readOGR(dsn = "/Users/marinaespinasse/github/nor-prep/prep/administrative/raw/shapefiles/NO_AdminOmrader")
komlist <-read.csv("/Users/marinaespinasse/github/nor-prep/prep/administrative/komlist.csv", sep = ";")
```

# Preparing data and map layers
## Filtering out municipallities that are not in the project and changing CRS
Here I also change the projection of the municipalities shape file.
```{r }
komlist <-filter(komlist, !Name == "NonCoast")
ohikom <-allkommuner[allkommuner$KOMM %in% komlist$Komnum,] 
ohifylke <-fylker[fylker$NAVN %in% c("Nordland", "Troms", "Finnmark"),]
kommuner.pol <- spTransform(ohikom, CRS("+proj=longlat +ellps=GRS80"))
fylker.pol <-spTransform(ohikom, CRS("+proj=longlat +ellps=GRS80"))
```

Load the world mapping layers and create an sf-object based on the municipalities shape file.
```{r }
world <- ne_countries(scale = "medium", returnclass = "sf")
kommunersf <- sf::st_as_sf(ohikom, coords= c("long", "lat"),crs = 4326, agr="identity")
fylkersf <-sf::st_as_sf(ohifylke, coords= c("long", "lat"),crs = 4326, agr="identity")
```

##  Mapping counties
```{r }
base_layer_counties <- ggplot() +
  geom_sf(
   data = world, 
   fill = "antiquewhite1", 
   color = "gray46", 
   size = 0.1) +
  geom_sf(data = fylkersf, fill = NA, aes(color = NAVN), size = 0.6) +
  coord_sf(xlim = c(10, 32), ylim = c(64,72)) +
  theme(axis.text = element_text(size = 17),
        legend.text = element_text(size = 15),
        legend.title = element_text(size = 15),
        legend.position = "none",
        panel.background = element_rect(fill = "aliceblue") ,
        panel.grid.major = element_line(color = "#dbdbd9", linetype = "dashed"),
        plot.title = element_text(hjust = 0.5, vjust = 0.5, size = 17, face = "bold")
  ) +
  scale_color_manual(values = c("#cc9351", "#8c3688", "#388f40" ))
 
base_layer_counties
```

```{r}
ggsave(paste0("figs/counties_map.pdf"), base_layer_counties, dpi = 300)
```


## Mapping municipalities with color based on county
First adding county column to kommunersf dataframe
```{r}
kommunersf_cny <- kommunersf %>% 
  mutate(county = case_when(
 KOMM >= 1800 &  KOMM < 1900 ~ "Nordland",
                                    KOMM >= 1900 &  KOMM < 2000 ~ "Troms",
                                    KOMM >= 2000 ~ "Finnmark"
 ))
```

```{r}
base_layer_mp <- ggplot() +
  geom_sf(
   data = world, 
   fill = "antiquewhite1", 
   color = "gray43", 
   size = 0.1) +
  geom_sf(data = kommunersf_cny, fill = NA, aes(color = county), size = 0.6) +
  coord_sf(xlim = c(10, 32), ylim = c(64,72)) +
  theme(axis.text = element_text(size = 17),
        legend.text = element_text(size = 15),
        legend.title = element_text(size = 15),
        legend.position = "none",
        panel.background = element_rect(fill = "aliceblue") ,
        panel.grid.major = element_line(color = "#dbdbd9", linetype = "dashed"),
        plot.title = element_text(hjust = 0.5, vjust = 0.5, size = 17, face = "bold")
  ) +
  scale_color_manual(values = c("#cc9351", "#8c3688", "#388f40" ))
 
base_layer_mp

```


```{r}
ggsave(paste0("./figs/municipalities_map.pdf"), base_layer_mp, dpi = 300)
```

```{r}
tiff(file = "./figs/norway_all_mcp.tiff",width = 7, height = 6, units = 'in', res = 300)
plot(allkommuner, border = grey(0.5)) 
dev.off()

```

